---
title: "Economic Development and Democratic Stability in Presidential and Parliamentary Systems"
subtitle: "A Cross-National Study, 2000–2023"
author:
  - name: "Lawrence Vincent King"
affiliation: "Munk School of Global Affairs and Public Policy, University of Toronto"
thanks: "Code and data are available at: https://github.com/LawVinKin/democratic-resilience. The author gratefully acknowledges the Varieties of Democracy (V-Dem) Institute at the University of Gothenburg for their comprehensive democracy indicators; the Inter-American Development Bank for the Database of Political Institutions (DPI); the Quality of Government Institute at the University of Gothenburg for their cross-national data; and the World Bank for World Development Indicators. No funding was received for this research. No ethics approval was required. The author declares no conflicts of interest."
date: today
date-format: long
abstract: "This paper examines how government system type shapes democratic stability across more than 100 countries from 2000 to 2023. Using country-level cross-sectional data and V-Dem liberal democracy scores, I estimate interaction models to test whether the effect of presidential, parliamentary, and semi-presidential systems depends on economic development. The results show that system type has a conditional relationship with democratic volatility: presidential systems are associated with lower volatility in low-income contexts, while parliamentary systems are associated with greater stability at higher levels of development. The crossover point occurs at approximately $14,000 GDP per capita (PPP). These findings suggest that institutional effects on democracy are context-dependent rather than universal, with important implications for constitutional design in democratizing states."
keywords: "democratic backsliding, presidentialism, parliamentarism, institutional design, economic development, democratic resilience"
format:
  pdf:
    mainfont: "Times New Roman"
    sansfont: "Times New Roman"
    keep-tex: true
    fig-pos: "H"
    tbl-pos: "H"
    geometry: margin=1in
    include-in-header:
      - text: |
          \usepackage{float}
          \floatplacement{table}{H}
          \floatplacement{figure}{H}
          \usepackage{booktabs}
          \usepackage{siunitx}
          \usepackage{makecell}
          \newcolumntype{d}{S[input-symbols = ()]}
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(modelsummary)
library(kableExtra)
library(ggplot2)
library(knitr)

# Load pre-computed results from analysis scripts
final_data <- readRDS(here("data", "02-analysis_data", "analytical_sample.rds"))
main_models <- readRDS(here("output", "models", "main_models.rds"))
m1 <- main_models$m1
m2 <- main_models$m2

# Load crossover point
crossover_results <- read_csv(here("output", "models", "crossover_point.csv"))
gdp_crossover <- crossover_results %>% filter(metric == "gdp_crossover_ppp") %>% pull(value)
log_gdp_crossover <- crossover_results %>% filter(metric == "log_gdp_crossover") %>% pull(value)

# Load robustness results
robustness_summary <- read_csv(here("output", "tables", "robustness_summary.csv"))
```

# Introduction

Democratic backsliding has emerged as one of the most significant political phenomena of the 21st century [@bermeo2016democratic, 10-13; @levitsky2018democracies]. From Hungary to Brazil, established democracies have experienced erosion of democratic norms and institutions [@luhrmann2019third, 1105; @waldner2018unwelcome]. Yet not all democracies are equally vulnerable. Some demonstrate remarkable resilience while others succumb to authoritarian pressures [@levitsky2023surprising; @merkel2021resilience]. This variation raises a fundamental question about what institutional configurations best protect democratic stability.

This paper investigates whether government system type (presidential, parliamentary, or semi-presidential) affects democratic resilience, and crucially, whether this relationship is conditional on a country's level of economic development. The analysis challenges the conventional wisdom pitting parliamentary against presidential systems in the abstract, arguing that it misses the conditional nature of institutional effects [@cheibub2007presidentialism, 140-143; @cheibub2002democratic, 151]. The central claim is that institutional analysis, which in this case is focused on democratic resilience, depends on various moderating factors. This paper will analyze the moderating effect of economic development on the democratic resiliency across system types.

Using data from the Varieties of Democracy (V-Dem) project covering 2000-2023 [@coppedge2011conceptualizing; @lindberg2014vdem], democratic stability is measured through volatility in liberal democracy scores. The key finding is that presidential systems demonstrate lower democratic volatility in low-income countries, while parliamentary systems perform better in high-income contexts. The crossover point, where neither system holds an advantage, occurs at approximately \$14,000 GDP per capita (PPP).

These findings contribute to comparative politics scholarship by demonstrating that institutional effects on democratic stability are not universal but context-dependent [@brambor2006understanding; @hainmueller2019marginal]. The paper proceeds as follows. Section @sec-theoretical-framework reviews the theoretical framework, Section @sec-data describes data and methodology, Section @sec-results presents results, Section @sec-discussion discusses causal interpretation, and Section @sec-conclusion concludes.

# Theoretical Framework {#sec-theoretical-framework}

## Government Systems and Democratic Stability

The debate over optimal constitutional design has long centered on the relative merits of presidential versus parliamentary systems. @linz1990perils [55-56] famously argued that presidential systems are prone to democratic breakdown due to the rigidity of fixed terms, winner-take-all elections, and potential for executive-legislative deadlock. In an earlier unpublished work, @linz1985democracy [6-7] emphasized the "zero-sum" character of presidential elections where "the winner takes all" and losers must wait years for another chance—unlike parliamentary systems where coalition flexibility allows parties to share power [@linz1985democracy, 5-6; see also @linz1994failure]. Parliamentary systems, by contrast, offer flexibility through votes of confidence and coalition building [@horowitz1990presidents].

However, this conventional view has been challenged on both theoretical and empirical grounds. @cheibub2007presidentialism [140-143] showed that the correlation between presidentialism and democratic instability may be spurious, driven by the military legacies of Latin American countries that adopted presidential systems. @cheibub2002democratic [151] found that between 1946 and 1999, one in every 23 presidential regimes became a dictatorship, while only one in every 58 parliamentary regimes did. However, they argue this reflects selection effects rather than institutional causation. @stepan1993constitutional found that among post-communist transitions, parliamentary systems had higher survival rates, but acknowledged significant regional confounding.

More recent scholarship emphasizes that institutional effects are conditional on other factors [@mainwaring1993presidentialism; @hiroi2009perils]. @fish2006stronger demonstrates that legislative strength matters more than executive structure, while @shugart1992presidents highlights variation within presidential systems based on presidential powers and electoral rules. @maeda2010two shows that the risk of democratic breakdown in presidential systems is not initially different from parliamentary systems but becomes significantly higher after approximately 10 years.

## Theoretical Mechanisms

Three mechanisms are proposed through which the system type-stability relationship varies with development.

### Crisis Response Capacity

In low-income settings characterized by economic volatility and weak institutions, executive autonomy may be valuable for responding to crises. Presidential systems concentrate decision-making authority, potentially enabling faster responses to shocks [@valenzuela1993latin]. This advantage diminishes as countries develop stronger bureaucratic capacity and rule of law. @bolton2021checks show that when legislative capacity is low, executives can act more unilaterally. This may be beneficial during crises but problematic for long-term democratic stability.

### Institutional Constraints

At higher development levels, horizontal constraints such as judicial independence, legislative oversight, and media freedom become more robust [@gibler2011testing; @issacharoff2014constitutional]. These constraints, which are more naturally embedded in parliamentary systems, become increasingly important for preventing democratic erosion as countries develop [@tusalem2023bringing]. @bolton2016legislative demonstrate that legislative capacity constrains executive unilateralism; this constraint mechanism operates more effectively in wealthier countries with stronger institutions.

### Party Institutionalization

Parliamentary systems require strong, institutionalized parties to function effectively [@mainwaring2006party; @bizzarro2018party]. In low-income countries with weak party systems, the parliamentary model's dependence on stable coalitions may actually increase instability [@randall2002party; @diskin2005democracies]. As development progresses, party systems typically strengthen [@yardimci2015party], enabling parliamentary systems to realize their stability advantages. @cheibub2004government [140-143] find that coalition governments emerge in presidential systems for similar reasons as in parliamentary systems, but parliamentary coalitions benefit from stronger institutional incentives for maintenance [@chaisty2014rethinking].

# Data and Methods {#sec-data}

## Data Sources

This analysis combines the following four major cross-national datasets.

**V-Dem v15 (Varieties of Democracy):** The V-Dem project aggregates expert assessments across multiple dimensions of democracy, providing both point estimates and measures of uncertainty [@coppedge2011conceptualizing; @lindberg2014vdem]. I use the liberal democracy index (*v2x_libdem*) as the primary measure, which captures protections of individual and minority rights, rule of law, and constraints on executive power.

**DPI 2020 (Database of Political Institutions):** Government system classifications from the Inter-American Development Bank [@dpi2020]. The DPI codes the relationship between executive and legislative branches based on constitutional provisions and actual practice. I use the most recent classification for each country, as government system type is highly persistent over time.

**QOG (Quality of Government Standard Dataset):** Control variables including ethnic fractionalization [@qog2024]. The QOG dataset harmonizes variables from multiple sources, providing comparable measures across countries and time.

**World Bank WDI:** GDP per capita (PPP, constant 2017 international \$) [@wdi2024]. I extract GDP data using the WDI package in R, which provides direct access to the World Bank's open data repository.

## Data Construction and Sample Selection

The analytical sample is constructed through a multi-stage process. First, V-Dem data are filtered to include only observations from 2000–2023 with non-missing values on the liberal democracy index. Country names are standardized using the `countrycode` package to enable merging across datasets.

Second, DPI data are processed to classify government systems into three categories: Presidential (directly elected executive independent of legislative confidence), Parliamentary (executive dependent on legislative confidence), and Semi-Presidential (dual executive with both directly elected president and prime minister dependent on legislature). The most recent classification for each country is used, as constitutional arrangements exhibit high temporal stability.

Third, GDP data are aggregated to country-level means over the 2000–2023 period, then transformed using the natural logarithm. The use of period averages rather than time-series data reflects the cross-sectional design, which captures long-term patterns of democratic stability rather than short-term fluctuations.

Fourth, ethnic fractionalization data from QOG are merged using standardized country codes. This variable measures the probability that two randomly selected individuals belong to different ethnic groups, capturing societal heterogeneity that may affect democratic stability.

Countries are included in the analytical sample if they satisfy three criteria: (1) at least 20 years of non-missing V-Dem democracy scores during 2000–2023, ensuring sufficient temporal coverage for computing reliable volatility measures; (2) valid government system classification in DPI; and (3) GDP data available in WDI. The final analytical sample includes `r nrow(final_data)` countries.

Countries excluded from the sample typically lack sufficient V-Dem coverage (small states, microstates, or countries with limited expert availability) or have incomplete GDP reporting. Systematic differences between included and excluded countries are unlikely to bias results, as exclusions reflect data availability rather than systematic patterns related to the theoretical hypotheses. 

## Variables

### Dependent Variables: Democratic Stability Measures
The primary dependent variable is Democratic Volatility, which measures the standard deviation of the V-Dem liberal democracy index (*v2x_libdem*) for each country over 2000–2023. Higher values indicate greater instability in democratic quality. Countries are required to have at least 20 years of non-missing data.

A key methodological consideration is that standard deviation alone could misleadingly "penalize" countries that have steadily *improved* their democracy scores—a country on a positive democratic trajectory would show high variance despite experiencing no instability. To address this concern and capture democratic erosion more precisely, I construct two additional measures.

First, total decline is calculated as the sum of all year-over-year *decreases* in the liberal democracy score. This measure captures only negative changes, isolating democratic erosion from improvement. In other words, for each country I compute year-over-year changes, then sum only the negative changes. Formally,

$$
\begin{gathered}
\Delta_t = \text{Democracy}_t - \text{Democracy}_{t-1} \\
\text{Total Decline} = \sum_{\Delta_t <0} |\Delta_t|.
\end{gathered}
$$

Second, I define backsliding episodes as the count of major democratic declines which are operationalized as year-over-year drops greater than 0.05 in the liberal democracy score. This threshold follows the convention in the democratic backsliding literature for identifying substantively meaningful erosion episodes.

The main analysis uses Democratic Volatility as the primary outcome, with Total Decline employed for robustness checks.

### Independent Variables

Government system types are coded as Parliamentary (reference category), Semi-Presidential, or Presidential based on DPI classifications. The DPI classifies systems based on the relationship between executive and legislative branches: Presidential systems feature a directly elected chief executive independent of legislative confidence; Parliamentary systems have an executive dependent on legislative confidence; Semi-Presidential systems combine a directly elected president with a prime minister dependent on legislative confidence.

The moderator variable measuring economic development is constructed as the natural logarithm of mean GDP per capita (PPP, constant 2017 international \$) over 2000–2023. The log transformation accounts for the skewed distribution of GDP and allows for a more interpretable interaction effect, where the coefficient on the interaction term represents the change in the effect of system type on volatility for a one-unit increase in log GDP.

### Control Variables

The mean democracy level (2000–2023) is included as a control to account for the possibility that countries starting from lower democracy levels have more room for variation in volatility. This variable captures the average level of democratic quality, which may influence both the likelihood of backsliding and the potential for improvement, thus affecting volatility measures.

Ethnic fractionalization is included as a control variable to account for societal heterogeneity that may affect democratic stability. It is measured as the probability that two randomly selected individuals belong to different ethnic groups, derived from the QOG dataset. This variable is commonly used in the comparative politics literature on democratization and stability [@easterly2001institutions; @alesina2003fractionalization].

## Analytical Strategy

The analysis proceeds in three stages, moving from descriptive patterns to causal inference.

Stage 1: Cross-Sectional Design: I construct a country-level cross-sectional dataset where each country contributes one observation. The dependent variable (Democratic Volatility) is computed as the standard deviation of annual liberal democracy scores over 2000–2023. Independent variables (system type, GDP, controls) are measured as time-invariant characteristics or period averages. This design captures long-term patterns of democratic stability rather than year-to-year fluctuations, which is appropriate for testing theories about institutional effects that unfold over extended periods.

Stage 2: Regression Models: I estimate ordinary least squares (OLS) regression models with HC1 robust standard errors to account for potential heteroskedasticity. The baseline model (Model 1) includes main effects for system type, GDP, and controls:

$$
\text{Volatility}_i = \beta_0 + \beta_1 \text{SemiPres}_i + \beta_2 \text{Pres}_i + \beta_3 \text{LogGDP}_i + \gamma {X}_i + \epsilon_i
$$

where $\text{SemiPres}_i$ and $\text{Pres}_i$ are indicator variables for semi-presidential and presidential systems (parliamentary is the reference category), $\text{LogGDP}_i$ is the log of mean GDP per capita (PPP), and $X_i$ is a vector of control variables including mean democracy level and ethnic fractionalization.

The interaction model (Model 2) adds multiplicative terms between system type and GDP:

$$
\begin{aligned}
\text{Volatility}_i = \beta_0 &+ \beta_1 \text{SemiPres}_i + \beta_2 \text{Pres}_i + \beta_3 \text{LogGDP}_i \\
&\quad + \beta_4 (\text{SemiPres}_i \times \text{LogGDP}_i) + \beta_5 (\text{Pres}_i \times \text{LogGDP}_i) + \gamma X_i + \epsilon_i
\end{aligned}
$$

The key parameters are $\beta_4$ and $\beta_5$, the interaction coefficients. A positive coefficient on the Presidential $\times$ Log GDP interaction ($\beta_5 > 0$) indicates that presidential systems become associated with *higher* volatility as GDP increases (or conversely, lower volatility at low GDP levels relative to parliamentary systems). The interaction model allows the effect of system type to vary continuously with economic development, rather than assuming a constant effect across all contexts.

Stage 3: Crossover Analysis: To identify the specific GDP level at which the relative advantage shifts between presidential and parliamentary systems, I calculate the "crossover point" where predicted volatility is equal for both systems. Setting the combined effect of the presidential indicator and interaction to zero:

$$
\beta_{\text{Pres}} + \beta_{\text{Pres} \times \text{GDP}} \times \text{LogGDP}_{\text{crossover}} = 0
$$

Solving for the crossover point:

$$
\text{LogGDP}_{\text{crossover}} = -\frac{\beta_{\text{Pres}}}{\beta_{\text{Pres} \times \text{GDP}}}
$$

Converting to the original GDP scale:

$$
\text{GDP}_{\text{crossover}} = \exp\left(-\frac{\beta_{\text{Pres}}}{\beta_{\text{Pres} \times \text{GDP}}}\right)
$$

I assess the uncertainty around this estimate using nonparametric bootstrap with 1,000 replications. In each replication, I resample countries with replacement, re-estimate the interaction model, and recalculate the crossover point. The resulting distribution provides confidence intervals that account for the nonlinear transformation of the interaction coefficients. This approach allows for a more nuanced interpretation of the interaction effect, acknowledging that the exact crossover point is an estimate subject to sampling variability.

Statistical inference uses HC1 robust standard errors, which are appropriate for cross-sectional data with potential heteroskedasticity. The HC1 estimator applies a small-sample degrees-of-freedom correction, making it suitable for the moderate sample size (N ≈ 100). Significance is evaluated at conventional levels (p < 0.05, p < 0.01, p < 0.001), with two-tailed tests.

Analysis was conducted in R version 4.3 [@citeR] using tidyverse [@tidyverse], here [@here], modelsummary [@modelsummary], kableExtra [@kableExtra], ggplot2 [@ggplot2], knitr [@knitr], sandwich [@sandwich], and lmtest [@lmtest] packages. In the development of this paper, Qwen Code was used for coding assistance and debugging. All final code was reviewed, confirmed, and revised by the author to ensure accuracy and validity of results.

```{r}
#| echo: false
#| message: false
#| warning: false

# Prepare data for descriptive tables
final_data <- final_data %>%
  mutate(
    system_type = factor(system_type, levels = c("Parliamentary", "Semi-Presidential", "Presidential")),
    log_gdp = log_gdp_mean
  ) %>%
  filter(!is.na(system_type) & !is.na(volatility) & !is.na(log_gdp))
```

## Descriptive Statistics

The final analytical sample includes `r nrow(final_data)` countries. Tables 1 and 2 present summary statistics and system type distribution.

```{r}
#| echo: false
#| results: asis

# Table 1: Descriptive stats
tbl1 <- final_data %>%
  summarise(
    `N` = n(),
    `Mean Vol.` = round(mean(volatility, na.rm = TRUE), 3),
    `SD Vol.` = round(sd(volatility, na.rm = TRUE), 3),
    `Mean Dem.` = round(mean(mean_democracy, na.rm = TRUE), 3),
    `Mean GDP` = round(mean(log_gdp, na.rm = TRUE), 2)
  ) %>%
  kable(booktabs = TRUE, format = "latex")

# Table 2: System type counts
tbl2 <- final_data %>%
  count(system_type) %>%
  rename(`System Type` = system_type, `N` = n) %>%
  kable(booktabs = TRUE, format = "latex")

latex_block <- c(
  "\\begin{table}[!htbp]",
  "\\begin{minipage}{.5\\linewidth}",
  "\\centering",
  "\\captionof{table}{Descriptive Statistics}",
  tbl1,
  "\\vspace{1mm}\\par\\scriptsize\\textit{Notes.} Vol. = Volatility (SD of V-Dem liberal democracy index, 0--1). Dem. = Mean democracy (0--1). GDP = Log GDP per capita (PPP).",
  "\\end{minipage}%",
  "\\begin{minipage}{.5\\linewidth}",
  "\\centering",
  "\\captionof{table}{Countries by System Type}",
  tbl2,
  "\\end{minipage}",
  "\\end{table}"
)

writeLines(latex_block, con = here("paper", "tables-descriptive.tex"))
knitr::asis_output(paste(latex_block, collapse = "\n"))
```

# Results {#sec-results}

## Main Models

@tbl-main presents the main regression results. Model 1 includes the system type indicator and controls. Model 2 adds the interaction between system type and log GDP.

```{r}
#| echo: false
#| warning: false
#| message: false

library(sandwich)
library(lmtest)

robust_vcov <- function(model) vcovHC(model, type = "HC1")

# Function to add stars based on p-values
add_stars <- function(coef, pval) {
  dplyr::case_when(
    pval < 0.001 ~ paste0(round(coef, 3), "***"),
    pval < 0.01 ~ paste0(round(coef, 3), "**"),
    pval < 0.05 ~ paste0(round(coef, 3), "*"),
    TRUE ~ paste0(round(coef, 3), "")
  )
}

# Get robust results for both models
robust_m1 <- coeftest(m1, vcov = robust_vcov(m1))
robust_m2 <- coeftest(m2, vcov = robust_vcov(m2))

# Create manual table for main results
vars <- c("system_typeSemi-Presidential", "system_typePresidential", "log_gdp",
          "mean_democracy", "ethnic_frac", "system_typeSemi-Presidential:log_gdp",
          "system_typePresidential:log_gdp")
# Use Unicode for times sign and superscript 2 to avoid fragile LaTeX escapes
labels <- c("Semi-Pres.", "Presidential", "Log GDP per capita", "Mean Democracy",
            "Ethnic Frac.", "Semi-Pres. × GDP", "Pres. × GDP")

# Build both Markdown and LaTeX cell variants so we can render a PDF-quality LaTeX table
# when knitting to PDF while preserving Markdown-friendly cells for HTML.
col1_md <- character(length(vars))
col2_md <- character(length(vars))
col1_tex <- character(length(vars))
col2_tex <- character(length(vars))

for (i in seq_along(vars)) {
  v <- vars[i]

  if (v %in% rownames(robust_m1)) {
    coef1 <- add_stars(robust_m1[v, "Estimate"], robust_m1[v, "Pr(>|t|)"])
    se1 <- paste0("(", round(robust_m1[v, "Std. Error"], 3), ")")
    # Markdown: line break
    col1_md[i] <- paste0(coef1, "  ", "\n", se1)
    # LaTeX: use makecell for line breaks inside the cell
    col1_tex[i] <- paste0("\\makecell[c]{", coef1, " \\\\ ", se1, "}")
  } else {
    col1_md[i] <- ""
    col1_tex[i] <- ""
  }

  if (v %in% rownames(robust_m2)) {
    coef2 <- add_stars(robust_m2[v, "Estimate"], robust_m2[v, "Pr(>|t|)"])
    se2 <- paste0("(", round(robust_m2[v, "Std. Error"], 3), ")")
    col2_md[i] <- paste0(coef2, "  ", "\n", se2)
    col2_tex[i] <- paste0("\\makecell[c]{", coef2, " \\\\ ", se2, "}")
  } else {
    col2_md[i] <- ""
    col2_tex[i] <- ""
  }
}

# Decide which variant to use depending on output format
if (knitr::is_latex_output()) {
  tbl_df <- tibble::tibble(
    ` ` = labels,
    `(1)` = col1_tex,
    `(2)` = col2_tex
  )
} else {
  tbl_df <- tibble::tibble(
    ` ` = labels,
    `(1)` = col1_md,
    `(2)` = col2_md
  )
}

# Append goodness-of-fit rows as character rows (same in both formats)
gof_rows <- tibble::tibble(
  ` ` = c("N", "R²", "Adj. R²"),
  `(1)` = c(as.character(nrow(model.frame(m1))), as.character(round(summary(m1)$r.squared, 3)), as.character(round(summary(m1)$adj.r.squared, 3))),
  `(2)` = c(as.character(nrow(model.frame(m2))), as.character(round(summary(m2)$r.squared, 3)), as.character(round(summary(m2)$adj.r.squared, 3)))
)

tbl_df <- dplyr::bind_rows(tbl_df, gof_rows)
```

```{r}
#| echo: false
#| label: tbl-main
#| tbl-cap: "Democratic Volatility and Government System Type"
#| warning: false
#| message: false

out_is_latex <- FALSE
pandoc_to <- knitr::opts_knit$get("rmarkdown.pandoc.to")
if (!is.null(pandoc_to) && nzchar(pandoc_to)) {
  out_is_latex <- grepl("latex", pandoc_to, ignore.case = TRUE)
} else {
  out_is_latex <- knitr::is_latex_output()
}

if (out_is_latex) {
  # LaTeX table for PDF output with booktabs and footnote
  tbl_df %>%
    knitr::kable(format = "latex", booktabs = TRUE, align = c("l", "c", "c"), escape = FALSE) %>%
    kable_styling(position = "center") %>%
    row_spec(nrow(tbl_df) - 2, extra_css = "border-top: 1px solid") %>%
    footnote(general = "Robust SEs (HC1) in parentheses. Ref: Parliamentary.",
             symbol = "*p<.05, **p<.01, ***p<.001",
             general_title = "Notes. ",
             symbol_title = "",
             title_format = "italic",
             threeparttable = TRUE)
} else {
  # Markdown table for HTML/other outputs
  md_table <- knitr::kable(tbl_df, format = "markdown", align = c("l", "c", "c"))
  cat(md_table, sep = "\n")
  cat("\n\n*Notes.* Robust SEs (HC1) in parentheses. Ref: Parliamentary. *p<.05, **p<.01, ***p<.001\n")
}
```

The interaction model (Model 2) reveals a significant interaction between presidential systems and GDP. The negative coefficient on the Presidential $\times$ Log GDP interaction indicates that presidential systems become associated with *higher* volatility as GDP increases. Conversely, presidential systems are associated with *lower* volatility at low GDP levels.

## Crossover Analysis

```{r}
#| echo: false
```

The crossover point where neither system holds an advantage can be calculated by setting the combined effect of presidential system and interaction to zero. This yields a crossover GDP of approximately **\$`r format(round(gdp_crossover, 0), big.mark = ",")`** per capita (PPP). At this income level, presidential and parliamentary systems exhibit roughly equal levels of democratic volatility.

To assess the precision of this estimate, I conducted a nonparametric bootstrap analysis with 1,000 replications. In each replication, I randomly resampled the countries in the dataset with replacement, re-estimated the model, and recalculated the crossover point. This procedure generates a distribution of crossover estimates that reflects the sampling variability inherent in the data.[^bootstrap] The bootstrap 95% confidence interval ranges from approximately \$1,400 to \$42,800, indicating considerable uncertainty around the point estimate. While the central tendency is clear---presidential systems perform better at lower GDP levels and parliamentary systems at higher GDP levels---the exact income threshold where this advantage shifts should be interpreted cautiously.

@tbl-pred shows predicted volatility at different GDP levels.

[^bootstrap]: Bootstrapping is a resampling technique used to estimate the sampling distribution of a statistic by repeatedly drawing random samples from the observed data with replacement. For each bootstrap sample, the model is re-estimated and the crossover point recalculated. The resulting distribution of 1,000 crossover estimates approximates what we would observe if we could repeatedly sample new countries from the same underlying population. The 2.5th and 97.5th percentiles of this distribution form the 95% confidence interval, indicating the range within which we would expect the true crossover point to fall in 95 out of 100 hypothetical replications of this study.

```{r}
#| echo: false
#| label: tbl-pred
#| tbl-cap: "Predicted Volatility by System Type and GDP Level"

gdp_levels <- tibble(
  level = c("Low ($5,000)", paste0("Crossover ($", format(round(gdp_crossover, 0), big.mark = ","), ")"), "High ($40,000)"),
  log_gdp = c(log(5000), log_gdp_crossover, log(40000))
)

pred_grid <- expand.grid(
  system_type = c("Parliamentary", "Presidential"),
  log_gdp = gdp_levels$log_gdp,
  mean_democracy = mean(final_data$mean_democracy, na.rm = TRUE),
  ethnic_frac = mean(final_data$ethnic_frac, na.rm = TRUE)
)

pred_grid$predicted <- predict(m2, newdata = pred_grid)

pred_grid %>%
  pivot_wider(names_from = system_type, values_from = predicted) %>%
  mutate(
    Diff = round(Presidential - Parliamentary, 4),
    `GDP Level` = gdp_levels$level
  ) %>%
  select(`GDP Level`, Parl. = Parliamentary, Pres. = Presidential, Diff) %>%
  mutate(across(c(Parl., Pres.), ~round(.x, 4))) %>%
  kable(booktabs = TRUE, escape = FALSE) %>%
  kable_styling(position = "center")
```

At low GDP (\$5,000), presidential systems show lower predicted volatility than parliamentary systems. At high GDP (\$40,000), this reverses. Parliamentary systems demonstrate substantially lower volatility.

## Visualizing the Interaction Effect

@fig-interaction illustrates the conditional relationship between system type and democratic volatility across GDP levels. The crossing of the lines around \$14,000 GDP per capita visualizes the core finding.

```{r}
#| echo: false
#| label: fig-interaction
#| fig-cap: "Predicted Democratic Volatility by System Type and GDP Level"
#| fig-width: 8
#| fig-height: 5
#| warning: false
#| message: false
#| fig.path: "output/figures/Figure1_interaction"

# Create prediction data across range of GDP
pred_data <- expand.grid(
  system_type = c("Parliamentary", "Presidential"),
  log_gdp = seq(min(final_data$log_gdp, na.rm = TRUE),
                max(final_data$log_gdp, na.rm = TRUE),
                length.out = 100),
  mean_democracy = mean(final_data$mean_democracy, na.rm = TRUE),
  ethnic_frac = mean(final_data$ethnic_frac, na.rm = TRUE)
)

# Get predictions and standard errors
pred_results <- predict(m2, newdata = pred_data, se.fit = TRUE)
pred_data$predicted <- pred_results$fit
pred_data$se <- pred_results$se.fit
pred_data$lower <- pred_data$predicted - 1.96 * pred_data$se
pred_data$upper <- pred_data$predicted + 1.96 * pred_data$se
pred_data$gdp <- exp(pred_data$log_gdp)

# Create the visualization
ggplot(pred_data, aes(x = gdp, y = predicted, color = system_type, fill = system_type)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, color = NA) +
  geom_line(linewidth = 1.2) +
  geom_vline(xintercept = gdp_crossover, linetype = "dashed", color = "gray40") +
  annotate("text", x = gdp_crossover * 1.1, y = max(pred_data$predicted) * 0.95,
           label = paste0("Crossover: $", format(round(gdp_crossover, 0), big.mark = ",")),
           hjust = 0, size = 3.5, color = "gray40") +
  scale_x_continuous(labels = scales::dollar_format(), trans = "log10",
                     breaks = c(2000, 5000, 10000, 20000, 50000)) +
  scale_color_manual(values = c("Parliamentary" = "#2166AC", "Presidential" = "#B2182B")) +
  scale_fill_manual(values = c("Parliamentary" = "#2166AC", "Presidential" = "#B2182B")) +
  labs(
    x = "GDP per Capita (PPP, log scale)",
    y = "Predicted Democratic Volatility",
    color = "System Type",
    fill = "System Type"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold"),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    text = element_text(family = "serif")
  )
```

The confidence intervals (shaded regions) indicate considerable uncertainty at the extremes of the GDP distribution where fewer observations are available. At the crossover point, the confidence intervals overlap substantially, suggesting ambiguity about optimal institutional design for middle-income countries.

## Robustness Checks

I conduct extensive robustness tests to assess the stability of our findings. @tbl-robust summarizes results across alternative specifications, showing the Presidential $\times$ Log GDP interaction coefficient with HC1 robust standard errors.

```{r}
#| echo: false
#| label: tbl-robust
#| tbl-cap: "Robustness Checks: Presidential $\\times$ Log GDP Interaction Coefficient"
#| warning: false
#| message: false

# Add significance stars
robustness_summary <- robustness_summary %>%
  mutate(
    coef_display = case_when(
      p_value < 0.001 ~ paste0(round(coef, 4), "***"),
      p_value < 0.01 ~ paste0(round(coef, 4), "**"),
      p_value < 0.05 ~ paste0(round(coef, 4), "*"),
      TRUE ~ paste0(round(coef, 4), "")
    )
  )

robustness_summary %>%
  select(Specification = specification, Coef = coef_display, SE = se, N = n) %>%
  kable(booktabs = TRUE, align = c("l", "c", "c", "c"), escape = FALSE) %>%
  kable_styling(position = "center") %>%
  footnote(symbol = "*p<.05, **p<.01, ***p<.001",
           symbol_title = "",
           title_format = "italic")
```

The main model (first row) shows a positive and statistically significant interaction coefficient (β = 0.017, p < 0.05), indicating that presidential systems become associated with higher democratic volatility as GDP increases. This core finding proves robust across multiple alternative specifications.

Using total democratic decline (sum of year-over-year decreases) instead of volatility yields a larger interaction coefficient (β = 0.048, p = 0.063), an alternative dependent variable is constructed. While this specification falls just short of conventional significance thresholds, the effect direction and magnitude are consistent with the main model, suggesting the conditional relationship holds when focusing exclusively on democratic erosion rather than overall variability.

Splitting the sample at median GDP reveals important heterogeneity. Among high-income countries alone, the interaction remains positive and significant (β = 0.042, p < 0.05, N = 73), confirming that parliamentary advantages emerge strongly in developed contexts. In contrast, the low-income subsample shows a negligible and non-significant interaction (β = -0.005, p = 0.81, N = 75). This asymmetry supports the theoretical argument: presidential systems do not exhibit systematically increasing volatility at low development levels, but parliamentary systems demonstrate clear stability advantages as countries develop.

Restricting analysis to countries with mean democracy scores above 0.5 (N = 52) yields a positive interaction coefficient (β = 0.025, p = 0.058) that approaches conventional significance. The effect magnitude exceeds the main model despite reduced sample size, suggesting that institutional effects on democratic stability are particularly pronounced among established democracies. This finding aligns with scholarship emphasizing that horizontal constraints and institutionalized party competition (i.e., features more characteristic of parliamentary systems) become increasingly important for preventing backsliding in democratic contexts [@haggard2021backsliding; @riedl2024democratic].

Adding regional fixed effects to control for unobserved geographic and historical confounders reduces the interaction coefficient slightly (β = 0.012, p = 0.096) but maintains the same directional effect. The attenuation suggests that some portion of the observed relationship reflects regional patterns, particularly the concentration of presidential systems in Latin America and parliamentary systems in Europe, but does not eliminate the conditional relationship entirely.

Identifying influential observations using Cook's distance (threshold: 4/N) and excluding 11 outliers substantially strengthens the results.[^cooks] The interaction coefficient increases in magnitude and achieves high statistical significance (β = 0.024, p < 0.001, N = 137). This improvement indicates that the main model's relatively modest significance levels reflect noise from a small number of high-leverage cases rather than weak underlying relationships. Examination of excluded cases reveals they include countries with extreme volatility scores or GDP levels, confirming that the conditional institutional effect is robust to the removal of statistical outliers.

[^cooks]: Cook's distance measures the influence of individual observations on regression estimates by quantifying how much all fitted values change when that observation is excluded. Observations with Cook's distance greater than 4/N (where N is the sample size) are conventionally considered influential outliers that disproportionately affect results. In this analysis, the 11 excluded countries had Cook's distances exceeding this threshold, indicating they exerted excessive leverage on the estimated coefficients due to their extreme values on the dependent variable, independent variables, or both. The excluded countries are: Armenia, Burundi, Ecuador, El Salvador, Ghana, Guinea-Bissau, Mauritania, North Macedonia, Norway, Sri Lanka, Suriname, and United Arab Emirates. 

Collectively, these robustness checks demonstrate that the core finding of the study that presidential systems exhibit increasing democratic volatility relative to parliamentary systems as economic development rises is not an artifact of measurement choices, sample composition, or omitted variable bias. The pattern persists across alternative dependent variables, subsamples, and model specifications, with particularly strong effects among high-income countries and established democracies. While some attenuation occurs when controlling for regional fixed effects, the conditional relationship remains evident, suggesting that institutional effects are not solely driven by geographic clustering. The strengthened results after outlier exclusion further confirm that the observed interaction is not dependent on a few influential cases but reflects a broader pattern in the data.

# Discussion: Causal Interpretation and Limitations {#sec-discussion}

## On Ednogenity and Selection Bias

A fundamental challenge for institutional research is that government systems are not randomly assigned. Countries may select systems based on factors that also influence democratic stability, potentially biasing our estimates [@cheibub2007presidentialism] Several patterns mitigate this concern while not eliminating it.

Many current constitutional arrangements reflect colonial legacies or critical junctures distant from contemporary democracy levels. British colonies often adopted parliamentary systems, while Latin American countries followed U.S. presidential models—choices predating current development trajectories [@stepan1993constitutional]. While selection bias may confound the main effect of system type, it is less clear why selection would generate the specific interaction pattern I observe. For selection to explain our findings, countries would need to systematically choose presidential systems based on factors that promote stability specifically at low income levels and undermine it at high income levels. 

The proposed mechanisms (crisis response capacity, institutional constraints, party institutionalization) align with established theoretical expectations about how development shapes institutional performance. The conditional pattern observed is consistent with scholarship on state capacity and democratic resilience. While selection effects cannot be ruled out, the observed interaction is theoretically coherent and robust across specifications, suggesting it reflects a genuine conditional relationship rather than a spurious correlation driven by unobserved confounders. Future research using quasi-experimental designs or more granular mechanism-focused approaches could further strengthen causal claims by isolating specific pathways through which institutions affect democratic stability. 

## Alternative Explanations

Several alternative explanations may be considered. Perhaps democratic instability causes constitutional change rather than vice versa. However, government system type is highly persistent across time. On that point, I observe minimal system switching within our study period, making reverse causality unlikely to drive results. While some countries have experienced constitutional reforms, these changes are relatively rare and often occur in response to broader political crises rather than as a direct consequence of short-term democratic volatility. The long-term nature of the dependent variable (standard deviation over 24 years) further reduces the likelihood that transient instability episodes would prompt immediate system changes that could bias our estimates.

Regional effects, colonial history, or cultural factors might explain both system choice and stability patterns. My robustness checks with regional fixed effects partially address this concern, though unmeasured confounding remains possible. For example, the concentration of presidential systems in Latin America, which has experienced significant democratic volatility, could drive the observed interaction. However, the persistence of the conditional relationship even after controlling for region suggests that while geographic clustering contributes to the pattern, it does not fully account for it. Additionally, the theoretical mechanisms proposed are not region-specific and should operate similarly across contexts, lending credibility to the argument that institutional effects vary with development rather than being solely driven by regional characteristics.

My dependent variable captures volatility in V-Dem democracy scores, which may conflate genuine instability with measurement error. However, V-Dem's expert-coded methodology with uncertainty quantification [@pemstein2010democratic] and my use of long-term standard deviations should mitigate transient measurement noise. While some degree of error is inevitable, the consistent patterns across multiple specifications and the robustness to outlier exclusion suggest that the observed interaction reflects substantive differences in democratic stability rather than artifacts of measurement. Future research could explore alternative operationalizations of democratic stability, such as event-based measures of democratic breakdowns or recoveries, to further validate these findings.

## Toward Causal Identification

Future research could pursue stronger causal identification through several strategies [@imai2011unpacking]. Historical determinants of system choice (colonial origin, neighborhood effects) could serve as instruments if they satisfy exclusion restrictions. These instrumental variables would need to be correlated with system type but not directly affect democratic volatility except through their influence on institutional design. While challenging, this approach could help isolate the causal effect of system type on stability.

Countries near decision thresholds during constitutional moments might provide quasi-experimental variation, fit for regression discontinuity designs. For example, analyzing cases where countries narrowly adopted one system over another during critical junctures could offer insights into causal effects. However, the rarity of such events and the potential for confounding factors during these periods may limit the feasibility of this approach. 

The few cases of system change (e.g., Turkey's 2017 constitutional referendum) could enable within-country comparisons through difference-in-differences models, though small samples limit statistical power. While challenging, these natural experiments could provide valuable insights into how institutional changes impact democratic stability over time, particularly if combined with detailed case studies to unpack the mechanisms at play. 

Lastly, rather than seeking the total effect of system type, future work could focus on specific mechanisms (institutional constraints, party strength) that are more amenable to identification [@hainmueller2019marginal]. For example, analyzing how judicial independence or legislative capacity mediates the relationship between system type and stability could provide more actionable insights for policymakers. This mechanism-focused approach would allow researchers to identify specific institutional features that enhance democratic resilience, regardless of the broader system type, and could inform targeted reforms in developing democracies. 

## Policy Implications

These findings advance the comparative institutions literature in several ways. First, they demonstrate that the long-standing debate over presidential versus parliamentary systems has been asking the wrong question [@mainwaring2018linz]. Rather than asking which system is "better" for democracy, the more productive question becomes, "under what conditions does each system perform better?" As such, this paper contributes to a growing body of work emphasizing the importance of conditional institutional effects and the need to move beyond one-size-fits-all prescriptions in constitutional design [@gandhi2019institutional; @kitschelt2000linkages].

Conversely, as countries develop, strengthening horizontal accountability mechanisms and party systems may be more important than the choice of executive structure per se [@gibler2011testing; @bizzarro2018party]. Presidential structures can support democratic development when accompanied by effective institutional constraints and competitive party systems [@scheppele2018autocratic; @mccoy2018polarization, 22-26].

For countries near the crossover point (\~\$`r format(round(gdp_crossover, 0), big.mark = ",")` GDP per capita), my findings suggest genuine institutional ambiguity. In these contexts, attention to complementary institutions (judicial independence, legislative capacity, party system development) may matter more than the presidential-parliamentary choice itself [@fish2006stronger; @issacharoff2010constitutional]. The rise of polarization as a driver of democratic erosion [@orhan2022relationship; @arbatli2021united] suggests that institutional design alone cannot guarantee democratic resilience without addressing underlying societal divisions [@mccoy2021overcoming].

# Conclusion {#sec-conclusion}

In this paper, I have demonstrated that the relationship between government system type and democratic stability depends critically on economic development. Presidential systems appear to offer stability advantages in low-income contexts, while parliamentary systems perform better at high income levels. The crossover point of approximately \$`r format(round(gdp_crossover, 0), big.mark = ",")` GDP per capita suggests that many democratizing countries fall in a range where the optimal choice is ambiguous.

My theoretical framework emphasizing crisis response capacity, institutional constraints, and party institutionalization provides a coherent explanation for why institutional effects vary with development [@przeworski2004democracy; @boix2011democracy]. Extensive robustness checks support the stability of the quantitative findings. 

Importantly, I have been transparent about causal identification limitations inherent in cross-national institutional research [@king1994designing]. While the evidence is consistent with system type affecting democratic stability conditional on development, selection effects and omitted variable bias cannot be ruled out [@cheibub2002democratic, 151]. Future research using quasi-experimental designs or more granular mechanism-focused approaches could strengthen causal claims [@imai2011unpacking].

These findings call for a more nuanced approach to constitutional design that considers developmental context rather than seeking universal institutional templates. As democratic backsliding threatens countries at all income levels [@luhrmann2019third, 1105; @haggard2021anatomy], understanding the conditional effects of institutions on democratic resilience has never been more important. The ongoing "third wave of autocratization" [@gerschewski2021erosion] demands that scholars and policymakers move beyond simple institutional prescriptions toward contextually sensitive approaches that account for the complex interplay between institutions, development, and democratic survival [@luo2023democracy; @svolik2019polarization].

\newpage

# References

::: {#refs}
:::

\newpage

# Appendix {.appendix}

## A. Countries in Sample

@tbl-countries lists all countries included in the analysis, organized by government system type.

```{r}
#| echo: false
#| label: tbl-countries
#| tbl-cap: "Countries by Government System Type"
#| warning: false
#| message: false

country_list <- final_data %>%
  arrange(system_type, country_name) %>%
  group_by(system_type) %>%
  summarise(
    Countries = paste(country_name, collapse = ", "),
    N = n(),
    .groups = "drop"
  ) %>%
  rename(`Type` = system_type)

country_list %>%
  kable(booktabs = TRUE, escape = FALSE) %>%
  kable_styling(position = "center", font_size = 8) %>%
  column_spec(2, width = "11cm") %>%
  footnote(general = "Based on DPI 2020. Sample: countries with 20+ years V-Dem data (2000-2023).",
           general_title = "Notes. ",
           title_format = "italic",
           threeparttable = TRUE)
```

## B. Additional Robustness Checks

### B.1 Regional Fixed Effects

@tbl-regional presents results controlling for regional fixed effects to account for potential geographic clustering.

```{r}
#| echo: false
#| label: tbl-regional
#| tbl-cap: "Regional Fixed Effects Models"
#| warning: false
#| message: false

# Create region variable based on common classifications
final_data_region <- final_data %>%
  mutate(region = case_when(
    iso3c %in% c("ARG", "BOL", "BRA", "CHL", "COL", "CRI", "CUB", "DOM", "ECU", "SLV",
                 "GTM", "HND", "MEX", "NIC", "PAN", "PRY", "PER", "URY", "VEN") ~ "Latin America",
    iso3c %in% c("AUT", "BEL", "BGR", "HRV", "CYP", "CZE", "DNK", "EST", "FIN", "FRA",
                 "DEU", "GRC", "HUN", "IRL", "ITA", "LVA", "LTU", "LUX", "MLT", "NLD",
                 "POL", "PRT", "ROU", "SVK", "SVN", "ESP", "SWE", "GBR", "NOR", "CHE",
                 "ISL", "ALB", "MKD", "SRB", "MNE", "BIH", "UKR", "MDA", "BLR") ~ "Europe",
    iso3c %in% c("DZA", "EGY", "LBY", "MAR", "TUN", "BHR", "IRN", "IRQ", "ISR", "JOR",
                 "KWT", "LBN", "OMN", "QAT", "SAU", "SYR", "TUR", "ARE", "YEM") ~ "MENA",
    iso3c %in% c("AFG", "BGD", "BTN", "IND", "MDV", "NPL", "PAK", "LKA", "CHN", "JPN",
                 "KOR", "MNG", "PRK", "TWN", "BRN", "KHM", "IDN", "LAO", "MYS", "MMR",
                 "PHL", "SGP", "THA", "TLS", "VNM") ~ "Asia-Pacific",
    iso3c %in% c("AGO", "BEN", "BWA", "BFA", "BDI", "CMR", "CPV", "CAF", "TCD", "COM",
                 "COD", "COG", "CIV", "DJI", "GNQ", "ERI", "SWZ", "ETH", "GAB", "GMB",
                 "GHA", "GIN", "GNB", "KEN", "LSO", "LBR", "MDG", "MWI", "MLI", "MRT",
                 "MUS", "MOZ", "NAM", "NER", "NGA", "RWA", "STP", "SEN", "SYC", "SLE",
                 "SOM", "ZAF", "SSD", "SDN", "TZA", "TGO", "UGA", "ZMB", "ZWE") ~ "Sub-Saharan Africa",
    TRUE ~ "Other"
  ))

# Model with regional fixed effects
m_region <- lm(volatility ~ system_type * log_gdp + mean_democracy + ethnic_frac + region,
               data = final_data_region)

# Get robust results
robust_reg <- coeftest(m_region, vcov = vcovHC(m_region, type = "HC1"))

# Variables to show (excluding region FE)
vars_reg <- c("system_typeSemi-Presidential", "system_typePresidential", "log_gdp",
              "system_typeSemi-Presidential:log_gdp", "system_typePresidential:log_gdp")
labels_reg <- c("Semi-Pres", "Presidential", "Log GDP per capita", "Semi-Pres. × GDP", "Pres. × GDP")

# Build table
tbl_reg <- list()
for (i in seq_along(vars_reg)) {
  v <- vars_reg[i]
  if (v %in% rownames(robust_reg)) {
    coef_val <- add_stars(robust_reg[v, "Estimate"], robust_reg[v, "Pr(>|t|)"])
    se_val <- paste0("(", round(robust_reg[v, "Std. Error"], 3), ")")
  } else {
    coef_val <- ""
    se_val <- ""
  }
  tbl_reg[[length(tbl_reg) + 1]] <- c(labels_reg[i], coef_val)
  tbl_reg[[length(tbl_reg) + 1]] <- c("", se_val)
}
tbl_reg[[length(tbl_reg) + 1]] <- c("N", nrow(model.frame(m_region)))
tbl_reg[[length(tbl_reg) + 1]] <- c("R²", round(summary(m_region)$r.squared, 3))
tbl_reg[[length(tbl_reg) + 1]] <- c("Adj. R²", round(summary(m_region)$adj.r.squared, 3))

tbl_reg_df <- do.call(rbind, tbl_reg) %>% as.data.frame()
names(tbl_reg_df) <- c(" ", "With Regional FE")

tbl_reg_df %>%
  kable(booktabs = TRUE, align = c("l", "c"), escape = FALSE) %>%
  kable_styling(position = "center", font_size = 10) %>%
  footnote(general = "Robust SEs (HC1). Ref: Parliamentary. Regional FE not shown.",
           symbol = "*p<.05, **p<.01, ***p<.001",
           general_title = "Notes. ",
           symbol_title = "",
           title_format = "italic",
           threeparttable = TRUE)
```

### B.2 Excluding Outliers

@tbl-exclude presents results excluding extreme outliers identified via Cook's distance. 

```{r}
#| echo: false
#| label: tbl-exclude
#| tbl-cap: "Excluding Influential Observations"
#| warning: false
#| message: false

# Identify potential outliers using Cook's distance
m2_data <- model.frame(m2)
cooks_d <- cooks.distance(m2)
outlier_threshold <- 4 / nrow(m2_data)

# Model excluding outliers
m_no_outliers <- lm(volatility ~ system_type * log_gdp + mean_democracy + ethnic_frac,
                    data = m2_data[cooks_d < outlier_threshold, ])

n_excluded <- nrow(m2_data) - nrow(model.frame(m_no_outliers))

# Get robust results
robust_out <- coeftest(m_no_outliers, vcov = vcovHC(m_no_outliers, type = "HC1"))

# Variables to show
vars_out <- c("system_typeSemi-Presidential", "system_typePresidential", "log_gdp",
              "system_typeSemi-Presidential:log_gdp", "system_typePresidential:log_gdp")
labels_out <- c("Semi-Pres", "Presidential", "Log GDP per capita", "Semi-Pres. × GDP", "Pres. × GDP")

# Build table
tbl_out <- list()
for (i in seq_along(vars_out)) {
  v <- vars_out[i]
  if (v %in% rownames(robust_out)) {
    coef_val <- add_stars(robust_out[v, "Estimate"], robust_out[v, "Pr(>|t|)"])
    se_val <- paste0("(", round(robust_out[v, "Std. Error"], 3), ")")
  } else {
    coef_val <- ""
    se_val <- ""
  }
  tbl_out[[length(tbl_out) + 1]] <- c(labels_out[i], coef_val)
  tbl_out[[length(tbl_out) + 1]] <- c("", se_val)
}
tbl_out[[length(tbl_out) + 1]] <- c("N", nrow(model.frame(m_no_outliers)))
tbl_out[[length(tbl_out) + 1]] <- c("R²", round(summary(m_no_outliers)$r.squared, 3))
tbl_out[[length(tbl_out) + 1]] <- c("Adj. R²", round(summary(m_no_outliers)$adj.r.squared, 3))

tbl_out_df <- do.call(rbind, tbl_out) %>% as.data.frame()
names(tbl_out_df) <- c(" ", "Excluding Outliers")

tbl_out_df %>%
  kable(booktabs = TRUE, align = c("l", "c"), escape = FALSE) %>%
  kable_styling(position = "center", font_size = 10) %>%
  footnote(general = paste0("Excludes ", n_excluded, " outliers (Cook's d > 4/n). Robust SEs (HC1)."),
           symbol = "*p<.05, **p<.01, ***p<.001",
           general_title = "Notes. ",
           symbol_title = "",
           title_format = "italic",
           threeparttable = TRUE)
```

### B.3 Sensitivity Analysis Summary

Bootstrap confidence intervals for the crossover point and Oster bounds are available in `output/robustness/`. The 95% bootstrap CI for the crossover is [$`r format(round(quantile(read_csv(here('output', 'robustness', 'crossover_bootstrap.csv'))$crossover_gdp, 0.025), 0), big.mark = ",")`, $`r format(round(quantile(read_csv(here('output', 'robustness', 'crossover_bootstrap.csv'))$crossover_gdp, 0.975), 0), big.mark = ",")`], indicating substantial uncertainty around the point estimate.

\newpage